### 2018-05-23 Paper/examples.tex

We aim to construct a reduced-order surrogate for the following QoI:
\be
\mathcal{F}(\kappa, c, \theta) = \frac{1}{|D|} \int_D u(x; \kappa, c, \theta) \, dx, 
\label{eq:qoi}
\ee
%
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite-differences and Newton's method
was used to solve the resulting nonlinear system.~\alennote{Give the
number of degrees of freedom.} 
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left).~\alennote{Say how many PDE solves
were used.}
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an ROS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the ROS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the ROS in the present test problem
is expected to be much smaller. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_elliptic}
\caption{Left: Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample size for 
PCEs constructed in 10 and 5 dimensions to compare their convergence characteristics. 
Right: PDF of the QoI, $\mathcal{F}(\kappa, c, \theta)$ in~\eqref{eq:semilinear} is
plotted using the full-surrogate and the reduced-order PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}

>>>>

We aim to construct a reduced-order surrogate for the following QoI:
\be
\mathcal{F}(\kappa, c, \theta) = \frac{1}{|D|} \int_D u(x; \kappa, c, \theta) \, dx, 
\label{eq:qoi}
\ee
%
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite-differences, and Newton's method
was used to solve the resulting nonlinear system.~\alennote{Give the
number of degrees of freedom.} 
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left).~\alennote{Say how many PDE solves
were used.}
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an ROS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the ROS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the ROS in the present test problem
is expected to be much smaller. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_elliptic}
\caption{Left: Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample size for 
PCEs constructed in 10 and 5 dimensions to compare their convergence characteristics. 
Right: PDF of the QoI, $\mathcal{F}(\kappa, c, \theta)$ in~\eqref{eq:semilinear} is
plotted using the full-surrogate and the reduced-order PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}
